Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-05-15 manuscript/manuscript.Rmd
minor edits to discussion

We modeled each grass population based on two levels of data: individual and quadrat (Fig. 2). The individual data is the "raw" data. For the quadrat-level we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because previous analysis shows that measurement error at the individual-level is small [@Chu2014]. Based on these two datasets we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth (13 year-to-year transitions, 29 quadrats and 18,730 records across the four species), survival (13 years, 33 quadrats and 29,353 records across the four species), and recruitment (13 years, 33 quadrats and 304 records across the four species). At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1* (13 year-to-year transitions, 29 quadrats and 866 records across the four species). For modeling population growth at the quadrat level we ignore within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Given the relatively broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes.

@@

Our comparison of a traditional, demographic population model without environmental forcing (the IPM) to an equivalent model inspired by density-structured models (the QBM) showed that, generally, IPM forecasts of out-of-sample plant population states were no more accurate than forecasts from the QBM (Fig. 2; `no-climate' bars).
We expected the IPM to out-perform the QBM since the IPM includes more mechanistic detail on the perennial plant life cycle, but this was not the case.
Our comparisons of forecast accuracy relied on one-step-ahead forecasts, but we also found no evidence that the rate at which forecast accuracy decays as forecast horizon increases was different between models (Fig. 4).
Thus, it appears that, without environmental forcing, the IPM and QBM generate equivalent forecasts.
Such a finding confirms theoretical [@Freckleton2011] and empirical work [@Queenborough2011] showing that density-structured models can be useful surrogates for demographic models when the goal is to estimate or forecast population states over large spatial extents.
Likewise, the equivalency of forecast accuracy between the IPM and QBM held for all four species, when climate covariates are not included, extending the generality of previous findings [e.g., @Taylor2004; @Queenborough2011].

@@

The higher accuracy of the IPM with cliamte for *P. secunda*  highlights the superiority of contemporary model and variable selection approaches such as ridge regression and LASSO over techniques that lead to excluding "non-significant" effects from final models.
Especially in a Bayesian forecasting framework, ridge regression allows researchers to include forcing variables that we know are changing (e.g., temperature and precipitation) in final models even if the estimated effects are relatively small or even deemed unimportant due to noisy data, out of which it is more difficult to estimate effects on interannual variation.
Likewise, if a species is truly unresponsive to a given climate variable, statistical regularization techniques will shrink the mean and variance of a covariate estimate to zero [@Hooten2015].
No amount of statistical "machismo"" during the model fitting stage can make up for including the appropriate candidate covariates, which we attempted to do based on our knowledge of this semi-arid plant community.
However, the climate covariates we chose required aggregating daily weather data over specific time periods.
It may be that we chose the wrong time periods over which to aggregate.
New methods using functional linear models (or splines) may offer a data-driven approach for identifying the appropriate time periods over which to aggregate to produce a tractable set of candidate climate variables [@Teller2016].

>>>>

We modeled each grass population based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat-level we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because previous analysis shows that measurement error at the individual-level is small [@Chu2014]. Based on these two datasets of 13 year-to-year transitions, we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth, survival, and recruitment. At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1*. For modeling population growth at the quadrat level we ignore within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Sample sizes for each species and vital rate model are shown in Table 1. Given the relatively broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes.

@@

Our comparison of a traditional, demographic population model without environmental forcing (the IPM) to an equivalent model inspired by density-structured models (the QBM) showed that, generally, IPM forecasts of out-of-sample plant population states were no more accurate than forecasts from the QBM (Fig. 2; `no-climate' bars).
We expected the IPM to out-perform the QBM since the IPM includes more mechanistic detail on the perennial plant life cycle, but this was not the case.
Thus, it appears that, without environmental forcing, the IPM and QBM generate equivalent forecasts.
Such a finding confirms theoretical [@Freckleton2011] and empirical work [@Queenborough2011] showing that density-structured models can be useful surrogates for demographic models when the goal is to estimate or forecast population states over large spatial extents.
Likewise, the equivalency of forecast accuracy between the IPM and QBM held for all four species, when climate covariates are not included, extending the generality of previous findings [e.g., @Taylor2004; @Queenborough2011].

@@

The higher accuracy of the IPM with climate for *P. secunda*  highlights the superiority of contemporary model and variable selection approaches such as ridge regression and LASSO over techniques that lead to excluding "non-significant" effects from final models.
Especially in a Bayesian forecasting framework, ridge regression allows researchers to include forcing variables that we know are changing (e.g., temperature and precipitation) in final models even if the estimated effects are relatively small or even deemed unimportant due to noisy data, out of which it is more difficult to estimate effects on interannual variation.
Likewise, if a species is truly unresponsive to a given climate variable, statistical regularization techniques will shrink the mean and variance of a covariate estimate to zero [@Hooten2015].
No amount of statistical "machismo" during the model fitting stage can make up for including the appropriate candidate covariates, which we attempted to do based on our knowledge of this semi-arid plant community.
However, the climate covariates we chose required aggregating daily weather data over specific time periods.
It may be that we chose the wrong time periods over which to aggregate.
New methods using functional linear models (or splines) may offer a data-driven approach for identifying the appropriate time periods over which to aggregate to produce a tractable set of candidate climate variables [@Teller2016].
